---
title: "Course Project Practical Machine Learning"
author: "TK"
date: "15 décembre 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively.

These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks.

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 

In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 


## Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source:

http://groupware.les.inf.puc-rio.br/har. 

### About Data

#### Human Activity Recognition

Human Activity Recognition - HAR - has emerged as a key research area in the last years and is gaining increasing attention by the pervasive computing research community, especially for the development of context-aware systems. There are many potential applications for HAR, like: elderly monitoring, life log systems for monitoring energy expenditure and for supporting weight-loss programs, and digital assistants for weight lifting exercises.


#### Load packages

```{r results='hide', message=FALSE, warning=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(AppliedPredictiveModeling)
```



### Download data

```{r}
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "data_train.csv")

download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "data_test.csv")

```
### Read data

```{r}
training <- read.csv("data_train.csv", na.strings=c("NA","#DIV/0!",""))
validation <- read.csv("data_test.csv", na.strings=c("NA","#DIV/0!",""))
dim(training); dim(validation)
```

``` {r}
head(training$classe);class(training$classe); summary(training$classe)
```

## Cleaning Data

```{r}
### We Remove columns with Near Zero Values
NZV <- nearZeroVar(training, saveMetrics = TRUE)
myTraining <- training[, !NZV$nzv]

```

```{r}
### Remove columns with NA or is empty

myTraining <- myTraining[, names(myTraining)[sapply(myTraining, function (x)
        ! (any(is.na(x) | x == "")))]]

```

```{r}
### Removing  columns of the dataset that is unlikely to influence the prediction

Useless <-grepl("^X|timestamp|user_name", names(myTraining))
myTraining <- myTraining[, !Useless]
rm(Useless)

```
### Separate the data to be used for Cross Validation

Divide training set into 2 parts

```{r}
set.seed(1234)
inTrain <- createDataPartition(y=myTraining$classe, p=0.6, list=FALSE)
mytraining <- myTraining[inTrain, ]; mytesting <- myTraining[-inTrain, ]
dim(mytraining); dim(mytesting)

```

We can test 2 models: Decision Tree Model and Random forest

#### 1. Decision Tree Model

As the outcomes are categorical (nominal), a decision tree was the first model tested using the method rpart.

```{r}
set.seed(1234)
modFit1<-train(classe ~ .,method="rpart", data=mytraining)
print(modFit1$finalModel)

plot(modFit1$finalModel, main = "Classification tree")
text(modFit1$finalModel, use.n = TRUE, all = TRUE, cex = .7)
```{r}

```

```

```{r}

fancyRpartPlot(modFit1$finalModel,cex=.8,under.cex=0.4)

```

```{r}
predict1 <- predict(modFit1,mytesting)
RpartCM <-confusionMatrix(mytesting$classe, predict1)
RpartCM
```

#### 2. Random forest

```{r}
set.seed(1234)
modFit2 <- randomForest(classe ~ ., data=mytraining)
predict2 <- predict(modFit2, mytesting,  type = "class")
RFCM <-confusionMatrix(predict2, mytesting$classe)
RFCM
```


```{r}

plot(modFit2, lwd = 3, main = "Prediction model random forest")


```

#### Compare models and choose with the best accuracy (random forest)

```{r}
result <- data.frame(RpartCM$overall, RFCM$overall)
result
```

## Conclusion


Random Forest was a good model for prediction of exercise quality compared to rpart. The RF model had over 99% accuracy and fitted well to other subsamples of the data. 

In the first model D was the most difficult to predict. 


#### Generating Files to submit as answers for the Assignment:

We use Random Forests for our prediction:

```{r}
predictions <- predict(modFit2, validation, type = "class")
predictionsFinal <- data.frame(validation$problem_id, predictions)
print(predictionsFinal)

```

#### Function to generate files with predictions to submit for assignment

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictions)

```
